<!DOCTYPE html>




<html
 dir="ltr"
 lang="en"
 >
  <head>
    <meta charset="utf-8">
    <meta name="viewport" content="width=device-width, initial-scale=1">
    <meta name="theme-color" content=#ffffff>
    <link rel="stylesheet" href="/assets/css/app.css">
    <link rel="shortcut icon" type="image/png"
           href="/favicon.png" 
    />
    <script defer src="https://unpkg.com/alpinejs@3.9.0/dist/cdn.min.js"></script>
    <link rel="stylesheet" href="https://cdn.jsdelivr.net/npm/@fortawesome/fontawesome-free@5/css/all.min.css">
    
        <link rel="stylesheet" href="https://cdn.jsdelivr.net/npm/bulma-social@1/bin/bulma-social.min.css">
    
    <!-- Begin Jekyll SEO tag v2.8.0 -->
<title>Tutorials | BioDeepTime</title>
<meta name="generator" content="Jekyll v4.2.2" />
<meta property="og:title" content="Tutorials" />
<meta property="og:locale" content="en_US" />
<meta name="description" content="Write an awesome description for your new site here. You can edit this line in _config.yml. It will appear in your document head meta (for Google search results) and in your feed.xml site description." />
<meta property="og:description" content="Write an awesome description for your new site here. You can edit this line in _config.yml. It will appear in your document head meta (for Google search results) and in your feed.xml site description." />
<link rel="canonical" href="https://biodeeptime.github.io/tutorials/" />
<meta property="og:url" content="https://biodeeptime.github.io/tutorials/" />
<meta property="og:site_name" content="BioDeepTime" />
<meta property="og:type" content="website" />
<meta name="twitter:card" content="summary" />
<meta property="twitter:title" content="Tutorials" />
<script type="application/ld+json">
{"@context":"https://schema.org","@type":"WebPage","description":"Write an awesome description for your new site here. You can edit this line in _config.yml. It will appear in your document head meta (for Google search results) and in your feed.xml site description.","headline":"Tutorials","url":"https://biodeeptime.github.io/tutorials/"}</script>
<!-- End Jekyll SEO tag -->

    <!-- head scripts --></head>

  <body>
    <nav class="navbar is-primary " x-data="{ openNav: false }">
    <div class="container">
        <div class="navbar-brand">
            <a href="/" class="navbar-item">
                BioDeepTime
            </a>
            <a role="button" class="navbar-burger burger" aria-label="menu" aria-expanded="false" data-target="navMenu" :class="{ 'is-active': openNav }" x-on:click="openNav = !openNav">
                <span aria-hidden="true"></span>
                <span aria-hidden="true"></span>
                <span aria-hidden="true"></span>
            </a>
        </div>
        <div class="navbar-menu" id="navMenu" :class="{ 'is-active': openNav }">
            <div class="navbar-start">

                
                
                    
                    <a href="/about/" class="navbar-item ">About</a>
                    
                
                    
                    <div class="navbar-item has-dropdown is-hoverable ">
                        <a href="" class="navbar-link ">Database</a>
                        <div class="navbar-dropdown">
                            
                            <a href="/database/contributing/" class="navbar-item ">Contributors</a>
                            
                            <a href="/database/structure/" class="navbar-item ">Structure</a>
                            
                            <a href="/database/milestones/" class="navbar-item ">Milestones</a>
                            
                            <a href="/database/citation/" class="navbar-item ">Citation</a>
                            
                            <a href="/database/access/" class="navbar-item ">Access</a>
                            
                        </div>
                    </div>
                    
                
                    
                    <div class="navbar-item has-dropdown is-hoverable ">
                        <a href="" class="navbar-link ">Team</a>
                        <div class="navbar-dropdown">
                            
                            <a href="/team/people/" class="navbar-item ">People</a>
                            
                            <a href="/team/workshops/" class="navbar-item ">Workshops</a>
                            
                            <a href="/team/funding/" class="navbar-item ">Funding</a>
                            
                        </div>
                    </div>
                    
                
                    
                    <a href="/tutorials/" class="navbar-item is-active">Tutorials</a>
                    
                
                    
                    <a href="/publications/" class="navbar-item ">Publications</a>
                    
                
                    
                    <div class="navbar-item has-dropdown is-hoverable ">
                        <a href="" class="navbar-link ">Contribute!</a>
                        <div class="navbar-dropdown">
                            
                            <a href="/contribute/specs/" class="navbar-item ">Specifications</a>
                            
                            <a href="/contribute/contact/" class="navbar-item ">Workshops</a>
                            
                        </div>
                    </div>
                    
                
                
            </div>

            <div class="navbar-end">
                
            </div>

        </div>
    </div>
</nav>

    
        <section class="hero  is-medium  is-bold is-primary" >
    <div class="hero-body ">
        <div class="container">
            <h1 class="title is-2">Tutorials</h1>
            <p class="subtitle is-3"></p>
            
        </div>
    </div>
</section>
    
    


    <section class="section">
        <div class="container">
            <div class="columns is-multiline">
                
                <div class="column is-12">
                    
                    
                    
                    
                    
<div class="content">
    ```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(divDyn)
```

Add `toc: true` to the metadata of the .md file for a table of contents.

# Introduction

- Database was designed for use in research in mind
- Here are some tutorials on how to use it: 

* * *

# More tutorials

Additional tutorials will be made available on the [Evolv-ED](https://www.evolv-ed.net/) + logo 

* * * 

## Preparing data

Example R code chunks: 

```{r counts, message=FALSE, paged.print=FALSE}
# counts of specimens
counts <- c(35,19,13,9,6,4,2,2,2,2,2,2,2,1,1,1,1,1,1,1)
counts
```

We can have a citation here: such as [@kocsis2019package]

### Subheading

Some more R code

```{r specimens, message=FALSE, paged.print=FALSE}
# one letter that represents each species
species <- letters[1:length(counts)]

# extended format of the sample
specimens <- rep(species, counts)
```

Running the ``table()`` function confirms that the structure we created is indeed matching the count data above. 

```{r table, message=FALSE, paged.print=FALSE}
table(specimens)
```

To make the vector of specimens compatible with the ``subsample()`` function, we have to make it a ``data.frame``.

```{r df, message=FALSE, paged.print=FALSE}
samp <- data.frame(specimens, stringsAsFactors=FALSE)
```

### Subheading 2 

An example plot

```{r tabplot, fig.width=10}

plot(table(specimens))
```


# References


</div>
                </div>
                
            </div>
        </div>
    </section>
    
        <footer class="footer">
    <div class="container">
        
        

        <div class="content is-small has-text-centered">

        </div>
    </div>
</footer>

    
    <script src="/assets/js/app.js" type="text/javascript"></script><!-- footer scripts --></body>
</html>
