---
title: Tutorials
subttitle: Using BioDeepTime in analytical code
bibliography: 'bibliography.bib'
format:
  gfm: 
    variant: +yaml_metadata_block
permalink: /tutorials/
layout: page
---


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(divDyn)
```

Add `toc: true` to the metadata of the .md file for a table of contents.

# Introduction

- Database was designed for use in research in mind
- Here are some tutorials on how to use it: 

* * *

# More tutorials

Additional tutorials will be made available on the [Evolv-ED](https://www.evolv-ed.net/) + logo 

* * * 

## Preparing data

Example R code chunks: 

```{r counts, message=FALSE, paged.print=FALSE}
# counts of specimens
counts <- c(35,19,13,9,6,4,2,2,2,2,2,2,2,1,1,1,1,1,1,1)
counts
```

We can have a citation here: such as [@kocsis2019package]

### Subheading

Some more R code

```{r specimens, message=FALSE, paged.print=FALSE}
# one letter that represents each species
species <- letters[1:length(counts)]

# extended format of the sample
specimens <- rep(species, counts)
```

Running the ``table()`` function confirms that the structure we created is indeed matching the count data above. 

```{r table, message=FALSE, paged.print=FALSE}
table(specimens)
```

To make the vector of specimens compatible with the ``subsample()`` function, we have to make it a ``data.frame``.

```{r df, message=FALSE, paged.print=FALSE}
samp <- data.frame(specimens, stringsAsFactors=FALSE)
```

### Subheading 2 

An example plot

```{r tabplot, fig.width=10}

plot(table(specimens))
```


# References

